---
title: "Analytical Lecture on Multiple Regression"
author: "Jason Snyder"
date: "XX/XX/2021"
output:
  html_document: default
  pdf_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

We load the libraries we need for this script.  If your computer did not already install these packages, then you need to install them first.  To install packages, you can either: 

1. type in your **Console** window _**install.packages("here")**_, or 
2. select the **Tools** menu option followed by **Install Packages...**.

```{r libraries, message=FALSE}
### Load libaries for use in current working session
## Library "here" for workflow
library(here)

## tidyverse for data manipulation and plotting
# Loads eight different libraries simultaneously
library(tidyverse)

## broom to extract output cleanly from many statistical models including "lm"
library(broom) 

## knitr to produce pleasing tables
library(knitr) 

## plotly for 3D plotting
library(plotly)
```

## Load Data

We load the _**multiple_reg_driver.csv**_ data from the _**data**_ project directory folder.

```{r data}
## use read_csv() and here() to import the data file
data <- read_csv(
  ## use here() to locate file in our project directory
  here("data", "multiple_reg_driver.csv")
)
```

## Scatterplot in 3D

We examine the data with the **glimpse** function from **dplyr**.  Then, we produce a 3D scatterplot with the **plot_ly** from the **plotly** library.

```{r scatterplot}
### Examine data
glimpse(data)

### 3D plot without regression plane
## Select data and assign variables
scatter3d <- plot_ly(data = data, 
                     x = ~Interview, 
                     y = ~Cognitive, 
                     z = ~Performance, 
                     opacity = 0.6) %>% 
  ## Add data points  
  add_markers() %>% 
  ## Add axes labels
  layout(scene = list(
          xaxis = list(title = "Interview"),
          yaxis = list(title = "Cognitive"),
          zaxis = list(title = "Performance")))

## View the plot
scatter3d
```

## Multiple Regression

We estimate a multiple regression model where we regress **Performance** ratings onto drivers' scores on the **Interview** and **Cognitive** tests.

```{r model}
##### Multiple regression
#### Unstandardized coefficients
model_1 <- lm(Performance ~ Interview + Cognitive, data = data)

### Full summary
summary(model_1)

### Model summary
glance(model_1)

## Multiple correlation
sqrt(glance(model_1)$r.squared)

### Coefficients summary
tidy(model_1)

### Confidence intervals for single regression coefficients
## 95% confidence interval
# base
confint(model_1)

# tidy
confint_tidy(model_1)	

## 99% confidence interval
# base
confint(model_1, level = 0.99)

# tidy
confint_tidy(model_1, conf.level = .99)

## Standardized coefficients
model_1_z <- lm(scale(Performance) ~ scale(Interview) + scale(Cognitive), data = data)

# Full summary
summary(model_1_z)

# Model summary
glance(model_1_z)

# Coefficients summary
tidy(model_1_z)
```

### Fitted, Residual, and Predicted Values

We examine the fitted, residual, and predicted values from the unstandardized coefficients multiple regression model.

```{r fit_resid_pred}
### Base functions
## Fitted values for unstandardized model
fitted(model_1)

## Residual values for unstandardized model
resid(model_1)

### Using broom for fitted and residual values
## Save calculations into data object
data_aug <- augment(model_1, data)

## Examine specific rows and variables of the augmented data
# Select data
data_aug %>%
  # Choose specific rows
  slice(c(1:5, 20, 24:30, 33, 40, 45:50)) %>%
  # Select specific variables
  select(.fitted, .resid) %>%
  # Print the table
  kable(caption = "Augmented Data", digits = 2)

### Correlations
## Multiple correlation for fitted and observed outcome values
cor(data_aug$.fitted, data_aug$Performance)

## Coefficient of determination
cor(data_aug$.fitted, data_aug$Performance)^2

## Correlation between residuals and predictor, outcome, and fitted values
# Choose data
data_aug %>%
  # Select variables
  select(Interview, Cognitive, Performance, .fitted, .resid) %>%
  # Correlations
  cor()

### Predicting values on outcome for unobserved data
## Summary of data
summary(data)

## Choose predictor values and apply to regression equation
coef(model_1)[1] + coef(model_1)[2]*85 + coef(model_1)[3]*68
```

### Plotting Regression Plane

We plot a regression plane in the three-dimensional space of the variables in the multiple regression model.

```{r regression_plane}
### Plot the observed versus fitted outcome values
## Specify the data and mapping 
ggplot(data_aug, aes(x = .fitted, y = Performance)) + 
  ## Specify the geometry
  geom_point() + 
  #jitter layer to adjust points
  geom_jitter() +
  ## Specify the regression line
  geom_smooth(method = "lm", se = F) +
  ## Label the axes
  labs(x = "Fitted Values")

### Plotting regression plane
## Set-up predictor value grid
# Extract unique values for predictors
x_vals = seq(min(model_1$model$Interview),max(model_1$model$Interview),length.out = 50)
y_vals = seq(min(model_1$model$Cognitive),max(model_1$model$Cognitive),length.out = 50)

# Grid of predictor values
pred_data <- expand.grid(x_vals, y_vals)

# Name the columns of the prediction grid
names(pred_data) <- c("Interview", "Cognitive")

## Use regression model to generate prediction values
fit_vals <- predict(model_1, newdata = pred_data)

# Organize prediction values in a matrix
plane <- matrix(fit_vals, nrow = length(y_vals), 
                ncol = length(x_vals), byrow = T)

## Check transformations
# Regression plane
plane[3, 7]

# Fitted values
fit_vals[107]

# Regression equation
coef(model_1)[1] + coef(model_1)[2]*pred_data[107,1] + coef(model_1)[3]*pred_data[107,2]

## Call the 3D scatterplot
scatter3d %>% 
  ## Add regression plane to 3D scatterplot
  add_surface(x = ~ x_vals, y = ~ y_vals, z = ~ plane, showscale = T) %>%
  ## Add title for color scale
  colorbar(title = "Performance")
```

## Your Turn!

Two exercises to complete.

### Multiple Regression

Run a multiple regression with **Performance** and **Cognitive** as predictors of **Interview**.  

```{r your_turn_multiple_regression}

```

### Regression Plane

Produce a 3D scatterplot with a regression plane where **Performance** and **Cognitive** predict **Interview**.  

```{r your_turn_regression_plane}

```

## Model Comparison

We evaluate nested models to determine how much additional variance we can account for in **Performance** by **Cognitive** above and beyond **Interview**.

```{r model_comp}
### Restricted model
## Create model
model_2 <- lm(Performance ~ Interview, data = data)		

## Full summary
summary(model_2)

## Model summary
glance(model_2)

### Difference in R^2 between nested models
glance(model_1)$r.squared - glance(model_2)$r.squared

## Significance test
anova(model_2, model_1)
```
