---
title: "Simple Mediation"
author: "Jason Snyder"
date: "XX/XX/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

We load the libraries we need for this script.  If your computer did not already install these packages, then you need to install them first.  To install packages, you can either: 

1. type in your **Console** window _**install.packages("here")**_, or 
2. select the **Tools** menu option followed by **Install Packages...**.

```{r libraries, message=FALSE}
### Load libaries for use in current working session
## here for workflow
library(here)

## tidyverse for data manipulation and plotting
# Loads eight different libraries simultaneously
library(tidyverse)

## broom to extract output from statistical models
library(broom) 

## mediation for mediation models
library(mediation) 
```

## Load Functions

We load user-defined functions: **simpMedBoot** and **simpMedCarlo**.  These functions execute bootstrapping and Monte Carlo estimation for simple mediation models.

```{r functions}
### Source functions
source(here("scripts", "simple_mediation_functions.R"))
```

## Load Data

We load _**students.csv**_ from the _**data**_ project directory folder.

```{r data}
## Load data via the here function
students <- read_csv(
  (here("data", "students.csv"))
  )
```

## Baron and Kenny Approach

We evaluate the following mediation hypothesis:

_The effect of conscientiousness on life satisfaction is mediated by procrastination._

```{r baronKenny}
### Create desired data
## Write new data object
sel_stud <- students %>%
  ## Select variables
  dplyr::select(ls.1, consc, proc) %>%
  ## Keep only complete cases
  na.omit()

### Estimate first regression model
## c path (total effect)
m_1 <- lm(ls.1 ~ consc, data = sel_stud)

### Estimate second regression model
## a path 
m_2 <- lm(proc ~ consc, data = sel_stud)

### Estimate third regression model
## b and c' paths 
m_3 <- lm(ls.1 ~ consc + proc, data = sel_stud)

### Evaluate coefficients
## c path is significant:
summary(m_1)

## a path is significant:
summary(m_2)

## c' is not significant, b is significant:
summary(m_3)

### Sobel test
## Save the coefficients
# a coefficient
a <- coef(m_2)[2]

# b coefficient
b <- coef(m_3)[3]

# Estimate of indirect effect
a*b

# standard error of "a" path
se_a <- summary(m_2)$coefficients[2,2]

# standard error of "b" path
se_b <- summary(m_3)$coefficients[3,2]

## Compute z-value
z_sobel <- a*b / sqrt(b^2*se_a^2 + a^2*se_b^2 + se_a^2*se_b^2)

# Is the z value beyond the critical values (here: for alpha = .05)?
z_sobel < qnorm(.025) | z_sobel > qnorm(.975)
```

## Mediation Package

We estimate coefficients for the aforementioned simple mediation model using the **mediation** package.

```{r mediation}
### Specify regression models for mediator and outcome
## Mediator model
med_mod <- lm(proc ~ consc, data = sel_stud)

## Outcome model
out_mod <- lm(ls.1 ~ consc + proc, data = sel_stud)

### Summaries
## Mediator model
summary(med_mod)

## Outcome model
summary(out_mod)

### Mediation effects using bootstrapping
output_med_boot = mediate(med_mod, out_mod, 
                          treat = "consc", mediator = "proc", 
                          boot = TRUE)

## Examine results
summary(output_med_boot)

# Plot results
plot(output_med_boot)

### Mediation effects using Monte Carlo
output_med_carlo <- mediate(med_mod, out_mod, 
                    treat = "consc", mediator = "proc")

## Examine results
summary(output_med_carlo)

# Plot results
plot(output_med_carlo)

```

## Your Turn!

Using the the different approaches, estimate the mediation effect of **proc** on the relationship between **neuro** and **sat.stress.1**

```{r yourTurn}

```

## Monte Carlo Estimation

We estimate _Monte Carlo_ coefficients for the aforementioned simple mediation model using the user-defined function: **simpMedCarlo**.

```{r monteCarlo}
#### One Monte Carlo estimate
### What are the coefficients and standard errors again? 
a
b
se_a
se_b

### Simulate normally distributed sampling distribution for a
## Set seed for reproducibility
set.seed(167)

## Draw samples
sample_a <- rnorm(200000, mean = a, sd = se_a)

### Simulate normally distributed sampling distribution for b
## Set seed for reproducibility
set.seed(167)

## Draw samples
sample_b <- rnorm(200000, mean = b, sd = se_b)

### Look at distributions
## a coefficient
ggplot(data.frame(sample_a), aes(x = sample_a)) +
  ## Histogram geometry
  geom_histogram(bins = 20)

## b coefficient
ggplot(data.frame(sample_b), aes(x = sample_b)) +
  ## Histogram geometry
  geom_histogram(bins = 20)

### Select random values for a and b and multiply them
## Set seed
set.seed(203)

## Calculate single indirect effect Monte Carlo estimate
ab_carlo <- sample(sample_a, 1) * sample(sample_b, 1)

## Alternative method for Monte Carlo indirect effect estimate
ab_carlo <- rnorm(1, mean = a, sd = se_a) * 
  rnorm(1, mean = b, sd = se_b)

#### Multiple Monte Carlo estimates
### Simulate one Monte Carlo estimate
## Set seed
set.seed(321)

## Calculate a single estimate
simpMedCarlo(a = coef(m_2)[2], b = coef(m_3)[3], cp = coef(m_3)[2], 
             se_a = summary(m_2)$coefficients[2,2], 
             se_b = summary(m_3)$coefficients[3,2], 
             se_cp = summary(m_3)$coefficients[2,2])

### Simulate multiple Monte Carlo estimates
## Set seed
set.seed(345)

## Produce estimates with replicate
output_carlo <- replicate(1000,
  ## Call function
  simpMedCarlo(a = coef(m_2)[2], 
               b = coef(m_3)[3], 
               cp = coef(m_3)[2], 
               se_a = summary(m_2)$coefficients[2,2], 
               se_b = summary(m_3)$coefficients[3,2], 
               se_cp = summary(m_3)$coefficients[2,2]), 
  simplify = F) %>%
  ## Bind rows
  bind_rows()

## What are the means of the distributions? 
carlo_mean <- colMeans(output_carlo)

## What are the standard deviations (standard errors) of the distributions?
carlo_sd <- map_df(output_carlo, sd)

### What are the values of the 2.5th and 97.5th percentiles? 
### 95% Confidence Interval
## Create data frame
carlo_ci_res <- as.data.frame(
  ## Call map and define anonymous function
  map(output_carlo, function (ndx) {
    ## Contents of anonymous function
    quantile(ndx, probs = c(0.025, 0.5, 0.975))}))

### Create density plot
## Call data and mapping
ggplot(output_carlo, aes(x = ab_carlo)) +
  ## Density plot
  geom_histogram(aes(y = ..density..), color = "black", 
                 fill = "white", bins = 40) +
  ## Density curve
  geom_density(fill = "#FF6666", alpha = 0.2) +
  ## Vertical lines for quantiles
  geom_vline(data = carlo_ci_res, 
             mapping = aes(xintercept = ab_carlo),
             color="blue", linetype="dashed", size=1) +
  ## Labels for vertical lines
  geom_text(data = carlo_ci_res, 
            mapping = aes(x = ab_carlo, y = 0.4, 
                          label = paste(row.names(carlo_ci_res), 
                                        "=", round(ab_carlo, 2))), 
            size = 4, angle = 90, vjust = -0.4, 
            hjust = 0, color = "purple")
```