---
title: "Simple Mediation assignment template"
author: "YOUR NAME HERE"
date: "XX/XX/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

We load the libraries we need for this script.  If your computer did not already install these packages, then you need to install them first.  To install packages, you can either: 

1. type in your **Console** window _**install.packages("here")**_, or 
2. select the **Tools** menu option followed by **Install Packages...**.

```{r libraries, message=FALSE}
### Load libaries for use in current working session
## here for workflow
library(here)

## tidyverse for data manipulation and plotting
# Loads eight different libraries simultaneously
library(tidyverse)

## broom to extract output from statistical models
library(broom) 

## mediation for mediation models
library(mediation) 
```

## Load Functions

We load user-defined functions: **simpMedBoot** and **simpMedCarlo**.  These functions execute bootstrapping and Monte Carlo estimation for simple mediation models.

```{r functions}
### Source functions
source(here("scripts", "simple_mediation_functions.R"))
```

## Load Data

Load _**students.csv**_ from the _**data**_ project directory folder.

```{r data}

```

## Baron and Kenny Approach

Evaluate the following mediation hypothesis using the Baron and Kenny Approach:

_The effect of neuroticism (**neuro**) on life satisfaction (**ls.1**) is mediated by mood (**gb.1**)._

Create a new object called *sel_stud* that consists of only the complete cases of the variables life satisfaction, neuroticism, and mood

Estimate the c path regression model where life satisfaction (**ls.1**) is predicted by neuroticism (**neuro**).  Name this model object *m_1*

Estimate the a path regression model where mood (**gb.1**) is predicted by neuroticism (**neuro**).  Name this model object *m_2*

Estimate the b and c' path regression model where life satisfaction (**ls.1**) is predicted by neuroticism (**neuro**) and mood (**gb.1**).  Name this model object *m_3*

Apply the _summary_ function to each of the model objects to inspect which paths in the mediation model are significant

```{r baronKenny}

```

## Mediation Package

Using Monte Carlo estimation, estimate coefficients for the aforementioned simple mediation model using the **mediation** package.

Set your computer's random seed to 345 (i.e., *set.seed(345)*)

Create a model object for the mediator model where mood (**gb.1**) is predicted by neuroticism (**neuro**).  Name the object med_mod

Create a model object for the outcome model where life satisfaction (**ls.1**) is predicted by neuroticism (**neuro**) and mood (**gb.1**).  Name the object out_mod

Apply the _summary_ function to both of the model objects to inspect which paths are significant

Next, estimate the mediation effect using **Monte Carlo** estimation using the _mediate_ function from the **mediation** package

Inside the _mediate_ function, specify *med_mod* and *out_mod* as the first and second inputs, *neuro* as the treat variable and *gb.1* as the mediator variable. Name the model object *output_med_carlo* 

Apply _summary_ and _plot_ to *output_med_carlo* to inspect and plot the mediation effects, respectively

```{r mediation}

```

## Monte Carlo Estimation

Create a plot for the estimated _Monte Carlo_ coefficients for the aforementioned simple mediation model using the user-defined function: **simpMedCarlo**.

Set your computer's random seed to 345 (i.e., *set.seed(345)*)

Use the _replicate_ function to produce 1000 estimates.  Also inside of _replicate_, call the _simpMedCarlo_ function and specify the a, b, cp, se_a, se_b, and se_cp coefficients (*hint*: you do not need to change any of these from what is included in the walkthrough script).  Set simplify = F and pipe into the _bind_rows_ function.  Name the object *output_carlo*.

Create an object called *carlo_ci_res* by creating a data frame using the _map_ function.  Inside of _map_ specify the model input as *output_carlo* and define the anonymous function as ndx (i.e., *function(ndx)*).  For the contents of the anonymous function, call _quantile_ and set the first input to ndx, and request the 2.5th, 50th, and 97.5th percentiles (i.e., probs = c(0.025, 0.5, 0.975)) 

Lastly, use *ggplot* to create a density plot of the indirect effect.  Inside of *ggplot*, set the first input to *output_carlo* and map *ab_carlo* to the aestethic for the x-axis.  

Add a new layer for a density plot that uses _geom_histogram_ with *..density..* mapped to the aesthetic for the y-axis.  Set the color to black, fill to white, and bins = 40

Add another layer for the density curve that uses _geom_density_.  Set the fill to the hexadecimal value #FF6666 and the alpha value to 0.2

Add vertical lines for the for the quantiles with _geom_vline_.  Specify the data as *carlo_ci_res* and the aesthetic mapping for the xintercept = ab_carlo.  Set the color to blue, the linetype to dashed, and size = 1

Add a final layer using _geom_text_ to specify the labels for the vertical lines.

*Hint*: All of the code for this ggplot is the same as the code for the plot in the walkthrough script

```{r monteCarlo}

```