---
title: "Moderated Regression with Continuous Predictors Assignment"
author: "YOUR NAME HERE"
date: "XX/XX/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

We load the libraries we need for this script.  If your computer did not already install these packages, then you need to install them first.  To install packages, you can either: 

1. type in your **Console** window _**install.packages("here")**_, or 
2. select the **Tools** menu option followed by **Install Packages...**.

```{r libraries, message=FALSE}
### Load libraries for use in current working session
## here for workflow
library(here)

## tidyverse for data manipulation and plotting
# Loads eight different libraries simultaneously
library(tidyverse)

## broom to extract output from statistical models
library(broom) 

## plotly for 3D plotting
library(plotly)

## olsrr for regression diagnostics
library(olsrr)

## interactions to prepare 2D plot
library(interactions) 

## jtools to examine model results
library(jtools)
```

## Load Data

Load the _**students.RData**_ from the _**data**_ project directory folder.

```{r data}
## Load data via the here function
load(here("data", "students.RData"))
```

## Regression Models

Estimate regression models where procrastination (**proc**) is a function of awakeness (**at.1**) and conscientiousness (**consc**).  

### Main Effects Regression Model

Estimate a main effects (i.e., additive) regression model and review the results.

```{r mainEffects}
### Estimate model
mod_main_eff <- lm(proc ~ at.1 + consc, data = students)

## Review results
summary(mod_main_eff)
```

### Interaction Effect Regression Model

Estimate a main effects plus an interaction effect regression model and review the results.

Use the *augment* function to create a new data object called **mod_int_eff_data** that contains the fitted and residual values of the model.  

```{r intEffect}

```

### Compare Models

Compare the main effects model and interaction effect model.

```{r compMods}

```

### Multicollinearity

Examine the extent of collinearity in the interaction effect model.

```{r collinear}

```

### Centered Predictors

Center your predictors to eliminate non-essential collinearity amongst predictors.

```{r center}

```

### Regression Model with Centered Predictors

Estimate an interaction effect regression model with centered predictors.  Examine the variance inflation factor to see if centering predictors eliminated non-essential collinearity

```{r intEffCent}

```

### Create 2D Plot

Use the **effects** package and **ggplot** to create a 2D plot. (use centered conscientiousness **consc_cent** as the x-axis variable (i.e., pred = consc_cent) and centered awakeness as the moderator variable (i.e., modx = at.1_cent))

```{r plot2d}

```

### Calculate Simple Slopes

Calculate the simple slopes between conscientiousness and procrastination for different levels of awakeness.

```{r simpleSlopes}

```