---
title: "Categorical Predictors and Moderated Regression w/ Categorical Predictors"
author: "Jason Snyder"
date: "XX/XX/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

We load the libraries we need for this script.  If your computer did not already install these packages, then you need to install them first.  To install packages, you can either: 

1. type in your **Console** window _**install.packages("here")**_, or 
2. select the **Tools** menu option followed by **Install Packages...**.

```{r libraries, message=FALSE}
### Load libaries for use in current working session
## here for workflow
library(here)

## tidyverse for data manipulation and plotting
# Loads eight different libraries simultaneously
library(tidyverse)

## broom to extract output from statistical models
library(broom) 

## interactions to prepare 2D plot
library(interactions) 

## jtools for printing clean results tables
library(jtools)

## effects for calculating model statistics
library(effects)
```

## Load Data

We load _**students.RData**_ from the _**data**_ project directory folder.

```{r data}
## Load data via the here function
load(here("data", "students.RData"))
```

## Dummy Coding

We examine the **class** of **prev.home** from **students**. We apply dummy coding to estimate a regression model to examine mean differences on **age** for the categories of **prev.home**.  

```{r dummyCoding}
### Check the class of variables
## Check the class of all variables
glimpse(students)

## Check the class of specific variable
class(students$prev.home)

### Coding scheme of a factor variable
## Check the current coding scheme 
contrasts(students$prev.home)

## Change referent category
students$prev.home <- fct_relevel(students$prev.home, "Berlin")

## Check if coding scheme changed
contrasts(students$prev.home)

### Estimate regression model
mod_dum_code <- lm(age ~ prev.home, data = students)

## Examine results
summ(mod_dum_code)

## Compare with descriptive statistics from data
# Choose data
students %>%
  # Grouping variable
  group_by(prev.home = fct_explicit_na(prev.home)) %>%
  # Summarize
  summarize(group_mean = mean(age, na.rm = TRUE))
```

## Effect Coding

We apply effect coding to **prev.home** and fit a regression model to examine mean differences on **age** as a function of the categories of **prev.home**.

```{r effectCoding}
### Examine current coding scheme
contrasts(students$prev.home)

### Effect coding
## Example
contr.sum(4)

### Convert coding scheme to from dummy coding to effects coding
contrasts(students$prev.home) <- contr.sum(4)

## Examine change to coding scheme
contrasts(students$prev.home)

### Estimate regression model
mod_eff_code <- lm(age ~ prev.home, data = students)

## Examine results
summ(mod_eff_code)

## Unweighted grand mean from data
# Choose data
students %>%
  # Grouping variable
  group_by(prev.home = fct_explicit_na(prev.home)) %>%
  # Summarize
  summarize(group_mean = mean(age, na.rm = TRUE)) %>%
  # Filter rows
  filter(prev.home != "(Missing)") %>%
  # Calculate unweighted grand mean
  summarize(unwgt_grand_mean = mean(group_mean))

## Weighted grand mean from data
mean(students$age, na.rm = TRUE)

## Mean of the reference group
coef(mod_eff_code)[1] - sum(coef(mod_eff_code)[2:4])
```

## Combining Continuous and Categorical Predictors

We examine a regression model where both a _continuous_ and categorical predictors. 

```{r wrangling}
### Examine current coding scheme for variable
contrasts(students$sex)

### Select complete data
## Create data
sel_stud <- students %>%
  # Select variables
  select(ls.1, neuro, sex) %>%
  # Complete cases only
  na.omit() 

### Mean center
## Overwrite data
sel_stud <- sel_stud %>%
  # Create centered variable
  mutate(neuro_cent = neuro - mean(neuro))

```

```{r intEff}
### Estimate main effects regression model
mod_add_eff = lm(ls.1 ~ neuro_cent + sex, data = sel_stud)

## Examine results
summ(mod_add_eff)

### Estimate regression model with interaction eff
mod_int_eff = lm(ls.1 ~ neuro_cent * sex, data = sel_stud)

## Examine results
summ(mod_int_eff)

# Change in R^2 
glance(mod_int_eff)$r.squared - glance(mod_add_eff)$r.squared

# F-test on change in R^2
anova(mod_add_eff, mod_int_eff)

### Augment data using interaction effect model
sel_stud <- augment(mod_int_eff, sel_stud)
```

## Interaction Plot

We plot the interaction present in this model and these variables.

```{r plots}
### Produce 2D interaction plot
## Specify model
interact_plot(mod_int_eff, 
  ## Specify x-axis variable
  pred = neuro_cent, 
  ## Specify moderator variable
  modx = sex,
  ## Include points
  plot.points = TRUE,
  ## Labels
  x.label = "Neuroticism", y.label = "Life Sat.", legend.main = "Sex",
  ## Colors
  colors = c("red", "blue")) +
  ## Add limits
  ylim(8, 35)
```

## Simple slopes

We calculate the simple slopes for the interaction effect

```{r ss}
### Calculate simple slopes
## Specify model
ss <- sim_slopes(mod_int_eff, 
  ## Specify x-axis variable
  pred = neuro_cent, 
  ## Specify moderator variable
  modx = sex)

## examine simple slopes 
ss

### Plot simple slopes
plot(ss)


### Alternative 2D interaction plot
## Choose data and mapping
ggplot(sel_stud, aes(x = neuro_cent, y = ls.1, color = sex)) + 
  ## Choose point geometry
  geom_point() + 
  ## Choose smooth geometry
  geom_smooth(method = 'lm', formula = y ~ x, 
              fullrange = F, se = T, 
              aes(fill = sex)) +
  ## Labels for plot components
  labs(x = "Neuro", y = "Life Sat.",
       fill = "Gender", color = "Gender") 

### Calculating predicted values for different combination of predictor values
mod_int_eff_pred <- as.data.frame(effect("neuro_cent:sex", mod_int_eff))

## Examine results
mod_int_eff_pred

### Plot adjusted means and confidence intervals
## Select data
mod_int_eff_pred %>%
  ## Choose values to plot
  filter(neuro_cent == -1) %>%
  ## Choose mapping
  ggplot(aes(x = sex, y = fit)) + 
    # Bar plot
    geom_bar(stat = "identity", aes(fill = sex)) +
    # Error bars
    geom_errorbar(aes(ymin = lower, ymax = upper), 
                  width = .2, position = position_dodge(.9)) +
    # Labels
    labs(x = "Gender", y = "Life Sat.", fill = "Gender")
```

## Your Turn!

Estimate an interaction effect regression model by using _previous home_ and _awakeness_ (**at.1**) as predictors of life satisfaction (**ls.1** from **students**).  Mean center the continuous predictor, create a 2D interaction plot, and review the simple slopes.

```{r yourTurn}

```