---
title: "Analytical Lecture on Moderated Regression with Continuous Predictors"
author: "Jason Snyder"
date: "XX/XX/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

We load the libraries we need for this script.  If your computer did not already install these packages, then you need to install them first.  To install packages, you can either: 

1. type in your **Console** window _**install.packages("here")**_, or 
2. select the **Tools** menu option followed by **Install Packages...**.

```{r libraries, message=FALSE}
### Load libaries for use in current working session
## here for workflow
library(here)

## tidyverse for data manipulation and plotting
# Loads eight different libraries simultaneously
library(tidyverse)

## broom to extract output from statistical models
library(broom) 

## plotly for 3D plotting
library(plotly)

## olsrr for regression diagnostics
library(olsrr)

## interactions to prepare 2D plot
library(interactions) 

## jtools to examine model results
library(jtools)
```

## Load Data

We load _**students.RData**_ from the _**data**_ project directory folder.

```{r data}
## Load data via the here function
load(here("data", "students.RData"))
```

## Regression Models

We estimate regression models where life satisfaction (**ls.1**) is a function of mood (**gb.1**) and awakeness (**at.1**).  

### Main Effects Regression Model

We estimate a main effects (i.e., additive) regression model.

```{r mainEffects}
### Estimate model
mod_main_eff <- lm(ls.1 ~ gb.1 + at.1, data = students)

## Review results
summary(mod_main_eff)
```

### Interaction Effect Regression Model

We estimate a main effects plus an interaction effect regression model.

```{r intEffect}
### Estimate a model
mod_int_eff <- lm(ls.1 ~ gb.1*at.1, data = students)

## Review results
summary(mod_int_eff)

## Model fitted and residual values
mod_int_eff_data <- augment(mod_int_eff, mod_int_eff$model)
```

### Compare Models

We compare the main effects and interaction effect models.

```{r compMods}
### Calculate R-squared difference
glance(mod_int_eff)$r.squared - glance(mod_main_eff)$r.squared

## Nested models comparison
anova(mod_main_eff, mod_int_eff)
```

### Create 3D Plot

We plot the regression plane in 3D space.

```{r regPlane, eval=FALSE}
### Build the 3D scatterplot
scatter3d <- plot_ly(data = mod_int_eff_data, x = ~gb.1, y = ~at.1, z = ~ls.1, 
                     opacity = 0.6) %>% 
  add_markers() %>% 
  layout(scene = list(
    xaxis = list(title = "Mood"),
    yaxis = list(title = "Awake"),
    zaxis = list(title = "LS")))

### Make predictions from model
## Extract unique values for predictors
x_vals = seq(min(mod_int_eff_data$gb.1),max(mod_int_eff_data$gb.1),length.out = 50)
y_vals = seq(min(mod_int_eff_data$at.1),max(mod_int_eff_data$at.1),length.out = 50)

## Grid of predictor values
pred_data <- expand.grid(x_vals, y_vals)

# Name the columns of prediction grid
names(pred_data) <- c("gb.1", "at.1")

## Use regression model to generate prediction values
fit_vals <- predict(mod_int_eff, newdata = pred_data)

# Organize prediction values in a matrix
plane <- matrix(fit_vals, nrow = length(y_vals), 
                ncol = length(x_vals), byrow = T)

## Check transformations
# Regression plane
plane[10, 17]

# Fitted values
fit_vals[467]

# Regression equation
coef(mod_int_eff)[1] + 
  coef(mod_int_eff)[2]*pred_data[467, 1] + 
  coef(mod_int_eff)[3]*pred_data[467, 2] + 
  coef(mod_int_eff)[4]*pred_data[467, 1]*pred_data[467, 2]

### Plot regression plane
## Start with scatterplot
scatter3d_plane <- scatter3d %>% 
  ## Add regression plane
  add_surface(x = ~x_vals, y = ~y_vals, z = ~plane, showscale = T) %>% 
  ## Title for color scale 
  colorbar(title = "LS")

# Display plot
scatter3d_plane
```


### Multicollinearity

We examine the extent of collinearity in the interaction effect model.

```{r collinear}
### Examine correlations
## Choose data
mod_int_eff_data %>%
  ## Select variables
  select(gb.1, at.1) %>%
  ## Calculate interaction variable
  mutate(gb_at_int = gb.1*at.1) %>%
  ## Correlate
  cor()

### Examine variance inflation factor
ols_coll_diag(mod_int_eff)$vif_t
```

### Centered Predictors

We center predictors to eliminate non-essential collinearity amongst predictors.

```{r center}
### Center predictors
## Overwrite data
mod_int_eff_cent_data <- mod_int_eff_data %>%
  ## Compute centered variables
  # Select variables
  mutate_at(vars(gb.1, at.1), 
            # Center variables
            list(cent = ~ . - mean(.))) %>%
  ## Rename ID variable
  rename(id = .rownames) %>%
  ## Select variables
  select(id, ls.1, gb.1_cent, at.1_cent)
```

### Regression Model with Centered Predictors

We estimate an interaction effect regression model with centered predictors.

```{r intEffCent}
### Estimate regression model
mod_int_eff_cent <- lm(ls.1 ~ gb.1_cent*at.1_cent, data = mod_int_eff_cent_data)

## Review results
summary(mod_int_eff_cent)

### Fitted and residual values
mod_int_eff_cent_data <- augment(mod_int_eff_cent, mod_int_eff_cent_data)

### Examine correlations
## Choose data
mod_int_eff_cent_data %>%
  ## Select variables
  select(gb.1_cent, at.1_cent) %>%
  ## Calculate interaction variable
  mutate(gb_at_int = gb.1_cent*at.1_cent) %>%
  ## Correlate
  cor()

### Examine variance inflation factor after centering predictors
ols_coll_diag(mod_int_eff_cent)$vif_t
```

### Create 2D Plot

We use the **effects** package and **ggplot** to create a 2D plot.

```{r plot2d}
### Examine results with summ() from jtools package
summ(mod_int_eff_cent)

### Produce 2D interaction plot
## Specify model
interact_plot(mod_int_eff_cent, 
  ## Specify x-axis variable
  pred = gb.1_cent, 
  ## Specify moderator variable
  modx = at.1_cent,
  ## Include points
  plot.points = TRUE,
  ## Labels
  x.label = "Mood", y.label = "Life Satisfaction", legend.main = "Awake",
  ## Colors
  colors = c("red", "blue", "green"))
```

### Calculate Simple Slopes

We calculate the simple slopes between mood and life satisfaction for different levels of awakeness.

```{r simpleSlopes}
### Calculate simple slopes
## Specify model
ss <- sim_slopes(mod_int_eff_cent, 
  ## Specify x-axis variable
  pred = gb.1_cent, 
  ## Specify moderator variable
  modx = at.1_cent)

### Examine simple slopes
ss

### Plot simple slopes
plot(ss)
```

## Your Turn!

Estimate an interaction effect regression model by using mood and awakeness as predictors of procrastination (**proc** from **students**).  Mean center the predictors, create a 2D interaction plot, and review the simple slopes.

```{r yourTurn}

```