---
title: "Moderated Mediation Template Script"
author: "YOUR NAME HERE"
date: "XX/XX/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

We load the libraries we need for this script.  If your computer did not already install these packages, then you need to install them first.  To install packages, you can either: 

1. type in your **Console** window _**install.packages("here")**_, or 
2. select the **Tools** menu option followed by **Install Packages...**.

```{r libraries, message=FALSE}
### Load libaries for use in current working session
## here for workflow
library(here)

## tidyverse for data manipulation and plotting
# Loads eight different libraries simultaneously
library(tidyverse)

## broom to extract output from statistical models
library(broom) 

## mediation for mediation models
library(mediation) 
```

## Load Data

Load _**students.RData**_ from the _**data**_ project directory folder.

```{r data}

```

## Wrangle Data

Wrangle the loaded data to prepare for analyses.

Create a new data object called *sel_stud* that contains only the complete cases on the variables for mood (*gb.1*), neuroticism (*neuro*), life satisfaction (*ls.1*), and sex (*sex*)

Center the continuous predictor variables (mood, neuroticism, & life satisfaction) using the *mutate_at* function and adding the variables back to the *sel_stud* data object

```{r dataWrangle}

```

## Moderated Mediation Model

Estimate a moderated mediation model where **gb.1_cent** mediates the relationship between **neuro_cent** and **ls.1_cent** and **sex** moderates the path from **neuro_cent** to **gb.1_cent**.

Create an mediator model object called *med_mod* where the *gb.1_cent* is predicted by *neuro_cent*, *sex*, and their *interaction term*

Create an outcome model object called *out_mod* where ls.1_cent is predicted by *neuro_cent*, *sex*, and their *interaction*, as well as *gb.1_cent*

Use _summary_ on both model objects to review the results of the regression models

```{r modMed}

```


## Estimate moderated mediation effects

Next, you will estimate the moderated mediation effects for the different levels of *sex*

Create an object called *output_f* by specifying *med_mod* and *out_mod* inside of the _mediate_ function.  Specify *neuro_cent* as the treatment variable (i.e., treat = neuro_cent) and *gb.1_cent* as the mediator variable (i.e., mediator = gb.1_cent).  Set *boot = F*, *sims = 1000*, and *boot.ci.type = "bca"*.
Specify the covariate as *female*

Create another object called *output_m* by specifying *med_mod* and *out_mod* inside of the _mediate_ function.  Specify *neuro_cent* as the treatment variable (i.e., treat = neuro_cent) and *gb.1_cent* as the mediator variable (i.e., mediator = gb.1_cent).  Set *boot = F*, *sims = 1000*, and *boot.ci.type = "bca"*.
Specify the covariate as *male*

Apply _summary_ to both *output_f* and *output_m* objects

Lastly, create an object called *output_fm* by specifying *med_mod* and *out_mod* inside of the _mediate_ function.  Specify *neuro_cent* as the treatment variable (i.e., treat = neuro_cent) and *gb.1_cent* as the mediator variable (i.e., mediator = gb.1_cent).  Set *boot = F*, *sims = 1000*, and *boot.ci.type = "bca"*.

Inside of the _test.modmed_ function, specify *output_fm* as the model and then specify female and male as covariates to test whether the indirect effect is significantly different between women and men

```{r modMedOut}

```
