---
title: "Moderated Mediation Walkthrough"
author: "Jason Snyder"
date: "XX/XX/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

We load the libraries we need for this script.  If your computer did not already install these packages, then you need to install them first.  To install packages, you can either: 

1. type in your **Console** window _**install.packages("here")**_, or 
2. select the **Tools** menu option followed by **Install Packages...**.

```{r libraries, message=FALSE}
### Load libaries for use in current working session
## here for workflow
library(here)

## tidyverse for data manipulation and plotting
# Loads eight different libraries simultaneously
library(tidyverse)

## broom to extract output from statistical models
library(broom) 

## mediation for mediation models
library(mediation) 
```

## Load Data

We load _**students.RData**_ from the _**data**_ project directory folder.

```{r data}
## Load data via the here function
load(here("data", "students.RData"))
```

## Wrangle Data

We wrangle the loaded data to prepare for analyses.

```{r dataWrangle}
### Create desired data
## Write new data object
sel_stud <- students %>%
  ## Select variables
  dplyr::select(ls.1, consc, proc, sex, age) %>%
  ## Keep only complete cases
  na.omit()

### Center continuous predictor variables
## Add variables to data object
sel_stud <- sel_stud %>%
  ## Create new variables
  # Select variables
  mutate_at(vars(ls.1, proc, consc, age), 
          # Create function
          list(cent = ~ . - mean(.)))
```

## Moderated Mediation Model 1

We estimate the moderated mediation model where **proc_cent** mediates the relationship between **consc_cent** and **ls.1_cent** and **consc_cent** moderates the path from **proc_cent** to **ls.1_cent**.

```{r modMed1}
### Mediator model
med_mod_1 <- lm(proc_cent ~ consc_cent, data = sel_stud)

### Outcome model
out_mod_1 <- lm(ls.1_cent ~ consc_cent*proc_cent, data = sel_stud)

### Examine results on original sample
## Mediator model
summary(med_mod_1)

## Outcome model
summary(out_mod_1)

### Estimate moderated mediation via bootstrap or Monte Carlo sampling
## Specify regression models
output_1 <- mediate(med_mod_1, out_mod_1, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  control.value = -sd(sel_stud$consc_cent), treat.value = sd(sel_stud$consc_cent))

## Examine results
summary(output_1)

## Plot results
plot(output_1)
```

## Moderated Mediation Model 2

We estimate the moderated mediation model where **proc_cent** mediates the relationship between **consc_cent** and **ls.1_cent** and **sex** moderates the path from **consc_cent** to **proc_cent**.

```{r modMed2}
### Mediator model
med_mod_2 <- lm(proc_cent ~ consc_cent*sex, data = sel_stud)

### Outcome model
out_mod_2 <- lm(ls.1_cent ~ consc_cent*sex + proc_cent, data = sel_stud)

### Examine results on original sample
## Mediator model
summary(med_mod_2)

## Outcome model
summary(out_mod_2)

### Estimate moderated mediation
## Model for females
output_2_f <- mediate(med_mod_2, out_mod_2, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  covariates = list(sex = "female"))

## Model for males
output_2_m <- mediate(med_mod_2, out_mod_2, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  covariates = list(sex = "male"))

## Examine results
# Model for females
summary(output_2_f)

# Model for males
summary(output_2_m)

### Test statistical significance of indirect effect between values on the moderator
## Set model without covariates specification
output_2 <- mediate(med_mod_2, out_mod_2, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca")

## Perform statistical test
# Specify model
test.modmed(output_2, 
  # Specify covariates lists
  covariates.1 = list(sex = "female"), 
  covariates.2 = list(sex = "male"))
```

## Moderated Mediation Model 3

We estimate the moderated mediation model where **proc_cent** mediates the relationship between **consc_cent** and **ls.1_cent** and **sex** moderates the path from **proc_cent** to **ls.1_cent**.

```{r modMed3}
### Mediator model: must include sex as a control even though it is not predicting proc in the path diagram
med_mod_3 <- lm(proc_cent ~ consc_cent + sex, data = sel_stud)

### Outcome model
out_mod_3 <- lm(ls.1_cent ~ consc_cent + proc_cent*sex, data = sel_stud)

### Examine results on original sample
## Mediator model
summary(med_mod_3)

## Outcome model
summary(out_mod_3)

### Estimate moderated mediation
## Model for females
output_3_f <- mediate(med_mod_3, out_mod_3, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  covariates = list(sex = "female"))

## Model for males
output_3_m <- mediate(med_mod_3, out_mod_3, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  covariates = list(sex = "male"))

## Examine results
# Model for females
summary(output_3_f)

# Model for males
summary(output_3_m)

### Test statistical significance of indirect effect between values on the moderator
## Set model without covariates specification
output_3 <- mediate(med_mod_3, out_mod_3, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca")

## Perform statistical test
# Specify model
test.modmed(output_3, 
  # Specify covariates lists
  covariates.1 = list(sex = "female"), 
  covariates.2 = list(sex = "male"))
```

## Moderated Mediation Model 4

We estimate the moderated mediation model where **proc_cent** mediates the relationship between **consc_cent** and **ls.1_cent** and **sex** moderates the path from **consc_cent** to **proc_cent** and **age_cent** moderates the path from **proc_cent** to **ls.1_cent**.

```{r modMed4}
### Mediator model: for the same reason as model 3, age is included even though it is not a predictor in the path diagram
med_mod_4 <- lm(proc_cent ~ consc_cent*sex + age_cent, data = sel_stud)

### Outcome model: Sex is also included in this model for the same reason
out_mod_4 <- lm(ls.1_cent ~ consc_cent*sex + proc_cent*age_cent, data = sel_stud)

### Examine results on original sample
## Mediator model
summary(med_mod_4)

## Outcome model
summary(out_mod_4)

### Estimate moderated mediation
## Model for females
output_4_f <- mediate(med_mod_4, out_mod_4, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  covariates = list(sex = "female"))

## Model for males
output_4_m <- mediate(med_mod_4, out_mod_4, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  covariates = list(sex = "male"))

## Model for younger
output_4_y <- mediate(med_mod_4, out_mod_4, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  covariates = list(age_cent = -sd(sel_stud$age_cent)))

## Model for older
output_4_o <- mediate(med_mod_4, out_mod_4, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  covariates = list(age_cent = sd(sel_stud$age_cent)))

## Model for female, younger
output_4_fy <- mediate(med_mod_4, out_mod_4, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  covariates = list(sex = "female", age_cent = -sd(sel_stud$age_cent)))

## Model for female, older
output_4_fo <- mediate(med_mod_4, out_mod_4, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  covariates = list(sex = "female", age_cent = sd(sel_stud$age_cent)))

## Model for male, younger
output_4_my <- mediate(med_mod_4, out_mod_4, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  covariates = list(sex = "male", age_cent = -sd(sel_stud$age_cent)))

## Model for male, older
output_4_mo <- mediate(med_mod_4, out_mod_4, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  covariates = list(sex = "male", age_cent = sd(sel_stud$age_cent)))

## Examine results
# Model for females
summary(output_4_f)

# Model for males
summary(output_4_m)

# Model for younger
summary(output_4_y)

# Model for older
summary(output_4_o)

# Model for female, younger
summary(output_4_fy)

# Model for female, older
summary(output_4_fo)

# Model for male, younger
summary(output_4_my)

# Model for male, older
summary(output_4_mo)

### Test statistical significance of indirect effect between values on the moderator
## Set model without covariates specification
output_4 <- mediate(med_mod_4, out_mod_4, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca")

## Perform statistical test
# Specify model
test.modmed(output_4, 
  # Specify covariates lists
  covariates.1 = list(sex = "female", age_cent = -sd(sel_stud$age_cent)), 
  covariates.2 = list(sex = "male", age_cent = sd(sel_stud$age_cent)))
```

## Moderated Mediation Model 5

We estimate the moderated mediation model where **proc_cent** mediates the relationship between **consc_cent** and **ls.1_cent** and **sex** moderates the path from **consc_cent** to **proc_cent** and **sex** moderates the path from **proc_cent** to **ls.1_cent**.

```{r modMed5}
### Mediator model
med_mod_5 <- lm(proc_cent ~ consc_cent*sex, data = sel_stud)

### Outcome model
out_mod_5 <- lm(ls.1_cent ~ consc_cent*sex + proc_cent*sex, data = sel_stud)

### Examine results on original sample
## Mediator model
summary(med_mod_5)

## Outcome model
summary(out_mod_5)

### Estimate moderated mediation
## Model for females
output_5_f <- mediate(med_mod_5, out_mod_5, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  covariates = list(sex = "female"))

## Model for males
output_5_m <- mediate(med_mod_5, out_mod_5, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca", 
  ## Specify moderator values
  covariates = list(sex = "male"))

## Examine results
# Model for females
summary(output_5_f)

# Model for males
summary(output_5_m)

### Test statistical significance of indirect effect between values on the moderator
## Set model without covariates specification
output_5 <- mediate(med_mod_5, out_mod_5, 
  ## Specify variables
  treat = "consc_cent", mediator = "proc_cent", 
  ## Specify simulation options
  boot = F, sims = 1000, boot.ci.type = "bca")

## Perform statistical test
# Specify model
test.modmed(output_5, 
  # Specify covariates lists
  covariates.1 = list(sex = "female"), 
  covariates.2 = list(sex = "male"))
```
