---
title: "Logistic Regression Walkthrough"
author: "Jason Snyder"
date: "XX/XX/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

We load the libraries we need for this script.  If your computer did not already install these packages, then you need to install them first.  To install packages, you can either: 

1. type in your **Console** window _**install.packages("here")**_, or 
2. select the **Tools** menu option followed by **Install Packages...**.

```{r libraries, message=FALSE}
### Load libaries for use in current working session
## here for workflow
library(here)

## tidyverse for data manipulation and plotting
# Loads eight different libraries simultaneously
library(tidyverse)

## broom to extract output from statistical models
library(broom) 

## aod for statistical tests
library(aod)

## descr for statistical tests
library(descr)
```

## Load Data

We load _**school.RData**_ from the _**data**_ project directory folder.

```{r data}
## Load data via the here function
load(here("data", "school.RData"))
```

## Example Probability Curves

We produce a set of example probability curves.

```{r probCurve}
### Set parameters
## Create grid
prob_ex <- expand.grid(b0 = seq(-2, 2, 2), 
  b1 = seq(-2, 2, 2), x = seq(-5, 5, 0.1)) %>%
  ## Convert to tibble
  as_tibble() %>%
  ## Calculate probabilities
  mutate(p = exp(b0 + b1*x)/(1 + exp(b0 + b1*x)))

### Plot
## Set data and mapping
ggplot(prob_ex, aes(x = x, y = p)) +
  ## Choose geometry
  geom_line() +
  ## Facet
  facet_grid(b1 ~ b0)
```

## Simple Logistic Regression

We examine the probability of _graduate school admittance_ as a function of _undergraduate GPA_.

```{r simpLog}
### Examine categorical outcome
## Levels
levels(school$admit)

## Contrasts
contrasts(school$admit)

### Estimate simple logistic regression model
simp_mod <- glm(admit ~ gpa, data = school, family = "binomial")

## Examine summary
summary(simp_mod)

## Examine coefficients
tidy(simp_mod)

## Examine overall model results
glance(simp_mod) 

### Predictions
## Logit predictions and diagnostics
school_simp_aug <- augment(simp_mod)

## Logit predictions
school$simp_logit <- predict(simp_mod)

## Odds ratio predictions
school$simp_odds <- exp(predict(simp_mod))

## Probability predictions
school$simp_prob <- fitted(simp_mod)

### Accuracy of predictions
## Name result and choose data
acc_simp_mod <- school %>%
  ## Summarize
  # True positives
  summarize(tp = sum(simp_prob >= 0.4 & admit == "Yes"),
            # True negatives
            tn = sum(simp_prob < 0.4 & admit == "No"),
            # False positives
            fp = sum(simp_prob >= 0.4 & admit == "No"),
            # False negatives
            fn = sum(simp_prob < 0.4 & admit == "Yes"))

## Accuracy computations
acc_simp_mod %>%
  # Overall accuracy
  summarize(overall = (tp + tn)/(tp + tn + fp + fn),
            # Positive accuracy
            positive = tp/(tp + fp),
            # Negative accuracy
            negative = tn/(tn + fn),
            # Sensitivity
            sensitivity = tp/(tp + fn),
            # Specificity
            specificity = tn/(tn + fp))

### Convert coefficients to odds ratios
exp(coef(simp_mod))

## convert odds ratio to percent multiplicative increase
exp(coef(simp_mod))/(1 + exp(coef(simp_mod)))

## Odds ratio confidence intervals
exp(confint(simp_mod))

### Statistical tests
## Likelihood ratio test for model
LogRegR2(simp_mod)

## Wald test for regression coefficients
wald.test(b = coef(simp_mod), Sigma = vcov(simp_mod), Terms = 2)

### Plot
## Choose data and mapping
ggplot(school, aes(x = gpa, y = as.numeric(admit) - 1)) +
  ## Point geometry
  geom_point(alpha = 0.5) +
  ## Smooth geometry
  geom_smooth(method = "glm", se = TRUE, 
              method.args = list(family = "binomial")) +
  ## Axes labels
  labs(x = "GPA", y = "Admit")
```

## Multiple Logistic Regression

We examine the probability of _graduate school admittance_ as a function of _undergraduate GPA_ and _school prestige_.

```{r multLog}
### Examine categorical predictor
## Levels
levels(school$rank)

## Contrasts
contrasts(school$rank)

## Cross-Tabulations
xtabs(~ admit + rank, data = school)

### Estimate multiple logistic regression model
mult_mod <- glm(admit ~ gpa + rank, data = school, family = "binomial")

## Examine results
summary(mult_mod)

## Augmented data
school_mult_aug <- augment(mult_mod, se_fit = TRUE)

### Compute predictions for each observation
## Logit predictions
school$mult_logit <- predict(mult_mod)

## Odds ratio predictions
school$mult_odds <- exp(predict(mult_mod))

## Probability predictions
school$mult_prob <- fitted(mult_mod)

### Accuracy of predictions
## Name result and choose data
acc_mult_mod <- school %>%
  ## Summarize
  # True positives
  summarize(tp = sum(mult_prob >= 0.5 & admit == "Yes"),
            # True negatives
            tn = sum(mult_prob < 0.5 & admit == "No"),
            # False positives
            fp = sum(mult_prob >= 0.5 & admit == "No"),
            # False negatives
            fn = sum(mult_prob < 0.5 & admit == "Yes"))

## Accuracy computations
acc_mult_mod %>%
  # Overall accuracy
  summarize(overall = (tp + tn)/(tp + tn + fp + fn),
            # Positive accuracy
            positive = tp/(tp + fp),
            # Negative accuracy
            negative = tn/(tn + fn),
            # Sensitivity
            sensitivity = tp/(tp + fn),
            # Specificity
            specificity = tn/(tn + fp))

### Convert coefficients to odds ratios
exp(coef(mult_mod))

## Odds ratio confidence intervals
exp(confint(mult_mod))

### Statistical tests
## Likelihood ratio test for model
LogRegR2(mult_mod)

## Wald test for categorical predictor
wald.test(b = coef(mult_mod), Sigma = vcov(mult_mod), Terms = 3:5)

## Hierarchical test of models
anova(simp_mod, mult_mod, test = "Chisq")

### Plot
## Calculate probability bands
# Choose data
school_mult_aug <- school_mult_aug %>%
  # Calculate probability
  mutate(prob = plogis(.fitted),
         # Calculate lower band probability
         prob_lower = plogis(.fitted - 1.96*.se.fit), 
         # Calculate upper band probability
         prob_upper = plogis(.fitted + 1.96*.se.fit))

## Choose data and mapping
ggplot(school_mult_aug, aes(x = gpa, y = prob)) +
  ## Ribbon geometry for bands
  geom_ribbon(aes(ymin = prob_lower, ymax = prob_upper, fill = rank), 
              alpha = 0.2) + 
  ## Line geometry for predictions
  geom_line(aes(color = rank), size = 2) +
  ## Labels
  labs(x = "GPA", y = "Probability", color = "Rank", fill = "Rank")
```

## Your Turn!

Estimate the multiple logistic regression model where _admittance to graduate school_ is a function of _undergraduate GPA_ and _GRE_.

```{r yourTurn}

```