---
title: "Multiple Regression LAB"
author: "YOUR NAME HERE"
date: "XX/XX/2021"
output:
  html_document: default
  pdf_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

We load the libraries we need for this script.  If your computer did not already install these packages, then you need to install them first.  To install packages, you can either: 

1. type in your **Console** window _**install.packages("here")**_, or 
2. select the **Tools** menu option followed by **Install Packages...**.

```{r libraries, message=FALSE}
### Load libaries for use in current working session
## Library "here" for workflow
library(here)

## tidyverse for data manipulation and plotting
# Loads eight different libraries simultaneously
library(tidyverse)

## broom to extract output cleanly from many statistical models including "lm"
library(broom) 

## knitr to produce pleasing tables
library(knitr) 

## plotly for 3D plotting
library(plotly)
```

## Load Data

Use the **read_csv** and **here** functions to read in the data from the appropriate location in the project directory.

```{r data}

```

## Scatterplot in 3D

Use the **glimpse** function from **dplyr** to examine the data.  Then, produce a 3D scatterplot to visualize how **Personality** and **Sitdriving** influence **Performance** ratings with the **plot_ly** function from the **plotly** library.

```{r scatterplot}

```

## Multiple Regression

Estimate a multiple regression model where **Performance** ratings are regressed onto drivers' scores on the **Personality** and **Sitdriving** tests.

Use *unstandardized* predictors

```{r model}

```

### Fitted, Residual, and Predicted Values

Examine the correlation between residuals and predictor, outcome, and fitted values from the unstandardized coefficients multiple regression model.

```{r fit_resid_pred}

```

### Plotting Regression Plane

Plot a regression plane in the three-dimensional space of the variables in the multiple regression model.

```{r regression_plane}

```

## Model Comparison

Evaluate the difference in variance explained between the model you just created and a new model where **Performance** is regressed only on **Personality**

Hint: The two models are nested models 

```{r model_comp}

```
