---
title: "Categorical Predictors and Moderated Regression w/ Categorical Predictors Template"
author: "YOUR NAME HERE"
date: "XX/XX/2021"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

We load the libraries we need for this script.  If your computer did not already install these packages, then you need to install them first.  To install packages, you can either: 

1. type in your **Console** window _**install.packages("here")**_, or 
2. select the **Tools** menu option followed by **Install Packages...**.

```{r libraries, message=FALSE}
### Load libaries for use in current working session
## here for workflow
library(here)

## tidyverse for data manipulation and plotting
# Loads eight different libraries simultaneously
library(tidyverse)

## broom to extract output from statistical models
library(broom) 

## interactions to prepare 2D plot
library(interactions) 

## jtools for printing clean results tables
library(jtools)

## effects for calculating model statistics
library(effects)
```

## Load Data

Load the _**students.RData**_ from the _**data**_ project directory folder.

```{r data}

```

## Dummy Coding

Check to make sure that the **sex** variable from the **students** data is dummy coded (i.e., female is assigned a value of 0 and male is assigned a value of 1).  

```{r dummyCoding}

```

## Combining Continuous and Categorical Predictors
  
Next you will estimate regression models where both **sex** and agreeableness (**agree**) are predictors of extraversion (**extra**). 

First create a new data object called *sel_stud* that consists of the variables *extra*, *agree*, and *sex*.  Be sure to include complete cases only

Next, overwrite the *sel_stud* data to create a new variable called *agree_cent* that is the mean centered version of agreeableness

click on *sel_stud* in the global environment to make sure that the new variable was created

```{r wrangling}

```

Now you will estimate two regression models, a main effects model and an interaction effect model, where extraversion (*extra*) is a function of agreeableness (*agree*) and *sex*

For both models, make sure to use centered agreeableness (*agree_cent*)

Inspect both models using either the summary() or summ() functions

Use the *glance()* function to compute the change in R^2 between the models and use the *anova* function to see if the difference in R^2 between the models is significant

Lastly, use the interaction effect model to augment the data. Save the result back into the *sel_stud* data

```{r intEff}

```

## Interaction Plot

Plot the interaction present in the model using the *interact_plot* function.  Specify *agree_cent* as the pred variable (i.e., pred = agree_cent) and *sex* as the modx variable (i.e., modx = sex)

Make sure to appropriately label your axes and set the limits on the y axis to fit the scale of extraversion (i.e., set the yaxis to *ylim(1, 5)*)

```{r plots}

```

## Simple slopes

Calculate the simple slopes for the interaction effect using the *sim_slopes()* function 

Specify *agree_cent* as the pred variable (i.e., pred = agree_cent) and *sex* as the modx variable (i.e., modx = sex)

```{r ss}

```
